---
title: "DataAnalysis"
author: "Hayden Wainwright"
date: "1/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Analysis of Species Limb Width and Limb Length


## Initial Analysis

I will start by first loading in any packages I will need. Specifically I will need to use dplyr, knitr and ggplot2.

```{r}
library(dplyr)
library(ggplot2)
library(knitr)

```

I will now load in the data from the measurmeent file as the object MData

```{r}
MData <- read.csv("Data/measurement.csv", row.names = 1)
```

I wanted to change the collumn names to a 3 letter short form of Spc for speices, LmL for Limb length, LmW for limb width, UnW for units of width, and UnL for units of length. 

```{r}
MData <- rename(MData, Spc =Species, LmW =Limb.Width, UnW =unitsW, LmL =Limb.Length, UnL =unitsL)
```




## Volume of Limbs

I also want to calcume the volume of the limbs. To do this I needed to correct the numerical variable for each species' limb width and length with the units assigned to it. This is down below using the conditional element selection function ifelse(). I then reassigned the units for the dimensions of the limbs to all be cm after the correction was made. 

```{r}
MData$LmW <- ifelse(MData$UnW == "cm", 1, 0.1)*MData$LmW

MData$LmL <- ifelse(MData$UnL == "cm", 1, 0.1)*MData$LmL

MData$UnW <- "cm"

MData$UnL <- "cm"

```


Once these corrections were complete I was able to find the area of the limbs by treating them as a cylinder with the width being the diameter of the cylider. 

```{r}
MData <- mutate(MData, Volume = pi*LmL*((LmW/2)^2))

MData
```
## Grouping Data by Plants and Animals
It was decided to group the species as either animal or plant. A collumn was added to denote this distinction and the data were then grouped by this distinction. 

```{r}
MData$Type <- c("Animal", "Animal", "Animal", "Animal","Animal","Animal","Animal", "Plant", "Plant", "Plant")

SpeciesType <- group_by(MData, Type)

```

I then wanted to see how the length, width, and volume of limbs varied between plants and animals studied. To this end, a table was created displaying the mean and standard deviation of the aforementioned traits. 

```{r}
TraitsByGroup <- summarize(SpeciesType, Limb.Length.AVG = mean(LmL), Limb.Length.SD = sd(LmL), Limb.Width.AVG = mean(LmW), Limb.Width.SD = sd(LmW), Volume.AVG = mean(Volume), Volume.SD = sd(Volume))

 kable(TraitsByGroup, caption = "Table 1.0 Average (AVG) and standard deiation (SD) of limb width, length, and volume for a series of species grouped by plants and animals")

```

## Plotting 

I wanted a basic plot of the limb width against limb length to identify any potential correlation between the two measurements. I wanted this to be a simple 

```{r}
Limb.Graph1 <- ggplot(data = MData, aes(x=LmL, y=LmW))+
  geom_point()+
  labs(x="Limb Length (cm)", y="Limb Width (cm)")+
  theme_classic()

print(Limb.Graph1)
```

Now, I am interested in only plotting the width and length that correspond to limbs with above-average volume. To this end, I first calculated the mean limb volume and seperated my data based on whether the volume for each row was above or below the mean. I then only plotted the data for points corresponding to above average volume. 

```{r}
Volume.AVG <- mean(MData$Volume)

Above.AVG.Volume <- filter(MData, Volume>Volume.AVG)

Above.AVG.Graph <- qplot(LmL, LmW, data = Above.AVG.Volume, xlab ="Limb Length (cm)", ylab = "Limb Width (cm)")+ theme_classic()
  
print(Above.AVG.Graph)

```

Finally, I would like a table that displays only the species and their respective limb volumes that correspond to a volume less than the average. To do this I filtered my data to only include species below the average volume and then created a table using the function kable. 


```{r}
Lower.AVG.Volume <- filter(MData, Volume < Volume.AVG)

Lower.AVG.Volume <- data.frame("Species" = Lower.AVG.Volume$Spc, "Volume" = Lower.AVG.Volume$Volume)

kable(Lower.AVG.Volume, align = 'c', caption = "Table 2. Species name and limb volume for species with below average limb volume")


```





