---
title: "DataGeneration"
author: "Hayden Wainwright"
date: "1/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Generation

Here I create a vector containing 10 species of interest to me.


```{r}
Species <- c("Ursus maritimus", "Chrysemys picta", "Falco perefrinus", "Salvenlinus alpinus", "Psedacris crucifer", "Pantherophis guttatus", "Cardinalis cardinalis", "Quercus alba", "Acer saccharum", "Rosa Gallicanae")

```
I have also created a series of limb width measurements from a random sample of a normal distribution with mean of 10 and standard deviation of 10. 
```{r}
Limb.Width <- rnorm(10, mean=10, sd=10)

Limb.Width<-abs(Limb.Width)
print(Limb.Width)
```
I further created a boolean vector for the units of these limb width measurements that is a string of 10 items either being mm or cm, with both having a 0.5 probability of being drawn. 


```{r}
unitsW <- sample(c("cm","mm"), 10, replace=TRUE, prob = c(0.5, 0.5))

```


Limb length is another factor I would like to consider in the species I have created so I will create another vector containing another random sample of 10 numbers from a normal distribution with mean 100 and standard deviation of 100. 

```{r}
Limb.Length <- rnorm(10, mean=100, sd=100)

Limb.Length<-abs(Limb.Length)
```

The limb length measurement will also need units so I create a vector for those this the same as for the width measurement.

```{r}
unitsL <- sample(c("cm","mm"), 10, replace=TRUE, prob = c(0.5, 0.5))
```

I now want to combine all the measurements for each of my ten species into one data frame called MyData


```{r}
MyData <- data.frame(Species, Limb.Width, unitsW, Limb.Length, unitsL)

str(MyData)
```

Finally, I would like to save this dataframe as a .csv file with the name measurements

```{r}
write.csv(MyData, 'Data/measurement.csv')
```


