---
title: "Regex"
author: "Hayden Wainwright"
date: "1/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## BIOL 812 Week 4 Assignment
#### Hayden Wainwright

### Introduction and Initial Object Assignment
In this assignment I will be using both dplyr and regular expression based commands to modify the species measurement data created in last week's assignment. TO start I will be loading in dplyr, ggplot, and knitr packages.

```{r}
library(dplyr)
library(ggplot2)
library(knitr)

```
I will now be importing my measurment.csv file as MData

```{r}
MData <- read.csv("Data/measurement.csv")
```


### Creating a New Species Collumn

I would like to create a collumn of the species names that contains only the first letter and a period based on the variables in my original species collumn. 

```{r}

MData <- mutate(MData, Sp = gsub("(\\w)\\w+(\\s\\w+)","\\1\\.\\2",Species))

print(MData)
```

### Creating a New Data Set

Now that I have created a new collumn with a shortened version of the specie names I am going to create a new data set that contains the shortened Sp collumn instead of the full name species collumn. I also want this collumn to be at the front of the data set, in the place where the old species collumn was. 

```{r}
MData2 <- select(MData, Sp, Limb.Width:unitsL)

print(MData2)


```

This completes the new dataset. I will now knit my file and push the files to GitHub. 








